Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-07-21 cover_letter/response_letter.tex

``The effect of the applied bias upon the thermodynamics can be rigorously corrected for using standard reweighting techniques \noter{Add some appropriate references here, maybe: https://www.springer.com/gp/book/9783540384472; [19] A. M.  Ferrenberg  and  R. H.  Swendsen,OptimizedMonte Carlo Data Analysis,Phys. Rev. Lett.63, 1195 (1989).[20] S. Kumar, D. Bouzida, R. H. Swendsen, P. A. Kollman, andJ. M. Rosenberg,The Weighted Histogram Analysis Methodfor Free-Energy Calculations on Biomolecules. 1. TheMethod,J. Comput. Chem.13, 1011 (1992).[21] E. Gallicchio, M. Andrec, A. K. Felts, and R. M. Levy,Temperature Weighted Histogram Analysis Method, Rep-lica Exchange, and Transition Paths,J. Phys. Chem. B109,6722 (2005).[22] M. Souaille and B. Roux,Extension to the WeightedHistogram  Analysis  Method:  Combining  UmbrellaSampling with Free Energy Calculations,Comput. Phys.Commun.135, 40 (2001).[23] M. R. Shirts and J. D. Chodera,Statistically Optimal Analy-sis of Samples from Multiple Equilibrium States,J. Chem.Phys.129, 124105 (2008).}. \rood{Rigorous elimination of the bias in the kinetics is critical for the construction of robust and reliable Markov state models that reflect the true system dynamics and are uncontaminated by any residual effects of the biasing potentials used to induce good sampling. A number of approaches to correct the kinetics are also available, including Girsanov reweighting, transition-based reweighting analysis (TRAM), dynamic histogram analysis method (DHAM), and their derivatives~\citep{Prinz2011OptimalDynamics, Chodera2011DynamicalTemperatures, Stelzl2017DynamicSimulations, Donati2017GirsanovModels, Donati2018GirsanovSimulations, Quer2018ANCOORDINATES} \noter{Add ref: I:10.1103/PhysRevX.4.041018}. The application of these methods under the conditions of high bias necessary for good sampling can, however, present challenges for numerical convergence.} A number of coarse-grained DNA force fields have been developed that enable direct observation of these events over microsecond time scales via unbiased coarse-grained molecular dynamics simulations, \citep{Romano2013DNADependence, Hinckley2013AnHybridization, Maciejczyk2014DNAModel, Markegard2015, Dans2016MultiscaleDNA} which, up to an acceleration factor associated with the smoothing of the underlying free energy landscape inherent to the coarse-graining procedure, can preserve a faithful model of the unbiased dynamics and associated pathways. These models have previously been used to study biological phenomena such as nucleosome dynamics\citep{Lequieu2016Tension-dependentUnwrapping, Lequieu2017InSliding} and transcription factor binding \cite{Terakawa2015P53Simulations, Tan2018DynamicData} as well as nanotechnology applications such as strand displacement \citep{Srinivas2013OnDisplacement, Haley2020DesignDisplacement} and DNA origami \citep{Snodin2019Coarse-grainedOrigami, Doye2020TheOrigami}. \rood{In this work, we choose to employ a coarse-grained model for DNA \citep{Hinckley2013AnHybridization} that is sufficiently inexpensive to enable the collection of sufficient volumes of unbiased simulation trajectories and adequately sample configurational space that we do not need to appeal to biasing strategies to enhance convergence nor apply any \textit{post hoc} corrections to the thermodynamics or kinetics.}''

(II) Our specification of a single strand concentration of 7 mol/L in Section 2.1.1 on \rood{p.~7} was an (embarrassing) typo -- the true concentration is actually 7 mmol/L (i.e., 7 mM). This is only 3.5$\times$ higher than the 2 mmol/L experimental concentration, and so is actually quite a good approximation for the experimental conditions. We regret our error, and have corrected this text to now read:

``A single pair of self-complementary sequences were placed in a cubic periodic box with side length 7.8 nm corresponding to a single-strand concentration of \rood{7 mM}. \rood{This concentration is only 3.5$\times$ larger than the 2 mM concentration employed in our experimental analyses (Section 2.2).}''

(III) Finally, we deliberately chose to explore the melting temperature as the state point at which the hybridized and dehybridized populations are in balance and sought to probe the mechanism and kinetics of the hybridization/dehybridization transitions under these equilibrium conditions corresponding to no net thermodynamic preference for either state (i.e., minimally driven conditions). A favorable side effect of this choice was indeed that we observe approximately equal numbers of hybridization and dehybridization events in a single (sufficiently long) simulation, whereas temperatures away from $T_\mathrm{melt}$ would favor one over the other. It is an interesting question to consider how the mechanism and kinetics of hybridization/dehybridization may change with temperature. Exploring small excursions either side of $T_\mathrm{melt}$ may still be possible using unbiased simulations, but we concur with the reviewer that biased sampling would be required to observe sufficient numbers of hybridization (dehybridization) events at large excursions above (below) $T_\mathrm{melt}$.

In response to this point, we have modified text in the terminal paragraph of the Conclusions on \rood{p.~35} observing that this may be an interesting point for follow-up study:

``Going forward, we will extend this work to discern more general trends in sequence-dependent hybridization/dehybridization for a wider range of oligomer sequences and motivate strategies for experimental comparisons. \rood{We also suggest that the MSM approach followed in this work, possible coupled with biased sampling and reweighting techniques \noter{Add refs to TRAM / DHAM or whatever it is that PyEmma uses by default}, may be well-suited to exposing changes in the hybridization/dehybridization mechanism and kinetics as a function of temperature.}''

@@

\textcolor{blue}{Very small points:
+ p16 line 3: "Although the 3SPN.2 model reproduces melting temperatures relatively well, we observed a systematic 4 K under-prediction relative to experiment and so we apply a universal (+4) K corrective calibration to our computational results". This sentence confused me at first. I presume you are just shifting your simulated kd slow curves to better match experiment, rather than shifting the data so that the melting temperatures agree. You may want to be more explicit as I read it as the latter at first. (Note one shouldn't expect the melting temperatures to agree as very different concentrations are being used in simulation and experiment, and one also has to correct for finite-size effects when estimating melting points from single duplex simulations (https://doi.org/10.1088/0953-8984/22/10/104102)}

\textbf{Author Reply:}   (I) We regret our lack of clarity in our choice of wording. Yes, we simply shifted the $k_d$ curves by a universal temperature shift, determined to be +4 K, that produced the optimal match with the experimental data.

(II) We regret our erroneous reporting of a 7 M rather than 7 mM concentration as detailed in our previous response. As we have now clarified, our simulations and experiments are conducted at similar concentrations (7mM and 2 mM, respectively). We agree that we still cannot expect perfect agreement in the melting temperatures due to the residual small  difference in concentrations, but likely this is small in comparison to discrepancies introduced by the approximations inherent in the coarse-grained molecular model.

(III) We greatly appreciate this comment and suggestion to apply a fluctuation correction to our data in order to correct for artifacts associated with finite-size effects following the approach detailed in Ref.~\noter{Add https://doi.org/10.1088/0953-8984/22/10/104102 to bibliography and cite here using citep.}. We have reanalyzed our data under this correction and this has pleasingly resulted in improved agreement between our simulations and experiment. \rood{Indeed, the agreement is so much improved that we no longer need to apply the +4 K calibration correction.} \noter{True or false? Edit as appropriate.} The revised figure is reproduced below.

\noter{Insert figure and caption here. Probably you also need some edits to the caption to (i) eliminate / change the +4 K correction, (ii) state that a fluctuation correction has been applied.}

In light of this modification, we have also added the following text to the Section \rood{X.X.X} on \rood{p.~XX}: \noter{Maybe append this to the paragraph on p.~15 ending ``...a function of time $f_{unfrayed}(t) = exp(-k_d^{fast}t)$.'' }

\rood{To correct for finite-size effects associated with the simulation of a single duplex within a periodic simulation box, we apply the fluctuation correction proposed by Ouldridge et al.\ \noter{Add ref} to correct our predictions for $f_\mathrm{hybridized}$ and $f_\mathrm{unfrayed}$ bulk yields. In each case, we correct the values observed in our simulations -- $f^\mathrm{sim}$ = $f_\mathrm{hybridized}^\mathrm{sim}$ or $f_\mathrm{unfrayed}^\mathrm{sim}$ -- to the bulk corrected result -- $f$ = $f_\mathrm{hybridized}$ and $f_\mathrm{unfrayed}$ -- using the following relations:}

\noter{Fix everything between the \noter{========} to use the notation in our paper, not Doye's paper, such that the text above makes sense.}

\noter{========}

%following calculates the dimer fraction in bulk solution from simulations performed in a periodic box. We apply this to the dissociation fraction in order to better compare with the experimental slow response. We compute these values from the dimer fraction (DD) and monomer fraction (D, D).

@@

    \begin{equation}\label{}
    \phi = \frac{\mathrm{probability (DD))}}{\mathrm{probability (D, D)}}
    \end{equation}

    \begin{equation}\label{}
	f_{\infty} = (1+\frac{1}{4\phi}) - \sqrt{(1+\frac{1}{4\phi})^2 - 1}
	\end{equation}
\noter{========}

\noter{If necessary to also change this.} We have also reworded the unclear wording regarding the temperature correction in Section \rood{X.X.X} on \rood{p.~XX} to now read:

\rood{XXX}

%\noteg{Add a small methods paragraph to describe the temp-dependent fits? I think this would also help address reviewer \#3 confusions about the various temps used }

%We have since incorporated the finite size correction for homodimers reported in  https://doi.org/10.1088/0953-8984/22/10/104102 \citep{Ouldridge2010ExtractingSimulations} when computing the slow dissociation fits. We find that the results show even better agreement with experiment, motivating us to update figure 1 with these data as shown below \ref{fig:conc_correction}. In addition, we have added the units and log scale changes proposed by reviewers \#1 and \#3, respectively.

%\begin{figure}[ht!]
%	\begin{center}
%        \includegraphics[width=\textwidth]{./revision_figures/both_response_log_4K_shift.png}
%        \caption{Updated figure 1 of the main text. We made three changes: i) adding units to both plots, ii) accounting for finite size effects when computing the slow response, and iii) displaying the slow response on a log axis. These were proposed by reviewers \#1, \#2, and \#3 respectively}
%        \label{fig:conc_correction}
%	\end{center}
%\end{figure}

%\begin{figure}[ht!]
%	\begin{center}
%        \includegraphics[width=\textwidth]{./revision_figures/both_response_log_no_correction_4K_shift.png}
%        \caption{Updated figure 1 of the main text. We added units to both plots and displayed the slow response on a log axis as proposed by reviewers \#1 and \#3, respectively \noteg{This is with no correction and the results do not look all the different (homodimer correction is less than hetero). I'm leaning towards just using this given that the temp-dependent simulations are not performed in equlibrium and the correction appears unncessary}}
%        \label{fig:conc_no_correction}
%	\end{center}
%\end{figure}

%The following was added to the methods section to describe the fluctuation correction:

%With regard to the longer simulations used to construct SRV-MSMs, we maintain the same empirical melting temperatures so as to maximize (de)hybridization events. Although selecting a slightly lower temperature for each sequence may better approximate the bulk melting temperature, we do not expect this small shift to have a substantial impact on the sequence-dependent kinetic landscapes we have reported. Furthermore, as the reviewer indicated we are already operating at a higher concentration than experiment and would not be able to recover the exact conditions given any temperature shift alone.

Finally, we have also edited the following text in Section 3.1 on \rood{p.~16} to clarify that the fluctuation correction is only applied for the purposes of the experimental comparison in Section 3.1 and is not propagated throughout the remainder of the paper:

``The computational time scales and melting temperatures reported in the remainder of the paper are not corrected \rood{by these calibration corrections or finite-size fluctuation corrections} since we are only interested in the relative trends in the behaviors of the four sequences.''

@@

\begin{shaded}
\textbf{Response to Reviewer \# 3}
\end{shaded}

\textcolor{blue}{Jones et al present a combined experimental and computational paper that probes the dynamics of the DNA hybridization/de-hybridization transition. Through a Markov state model analysis of coarse-grained simulation, the authors identify states that contribute to the process, and present experimental evidence to support the role of these states. In particular, repetitive sequences are observed in simulations to have a number of misaligned states that are both appreciably stable and contribute to hybridization and de-hybridization pathways as intermediates. The role of these states is diminished by base pairs that disrupt the repetitive sequence. In the special case of a sequence with a strong core and weaker extremities, metastable "frayed" states are also observed. The presence of these states is assessed experimentally trough the analysis of relaxation timescales observed in T-jump IR spectroscopy; fast relaxation is attributed to fraying equilibria, and a broader fast relaxation to an increased heterogeniety due to the presence of mis-aligned intermediates.}

\textcolor{blue}{The paper is generally well-written, and, I believe, represents a good attempt to tackle this important process. I personally have had a manuscript "desk-rejected" by JACS on the basis that it is "too focussed on DNA biotechnology", and "primarily of interest to RNA and DNA chemists". The same statements could be made about this paper. However, I disagree with the conclusion that such papers are not appropriate for JACS. I would therefore recommend publication, subject to satisfactory responses to the comments below.  }

\textbf{Author Reply:}   We are delighted by the reviewer's warm reception of our paper, and pleased that they find it to be appropriate for publication in JACS.

@@

\textbf{Author Reply:}   We regret our (embarrassing) typo -- the concentration is in fact 7 mM, not 7 M. The calculation is below and we have made the correction to Section 2.1.1 on \rood{p.~7}. We regret our error.

\begin{equation}\label{}
\frac{2 \mathrm{\:molecules}}{\mathrm{box\:volume}} = \frac{2 \mathrm{\:molecules}}{(7.80 \times 7.80 \times 7.80) \mathrm{\:nm}^3} = \frac{3.32 \times 10^{-24} \mathrm{\:moles}}{4.75 \times 10^{-25} \mathrm{\:m}^3} = 7.0 \frac{\mathrm{moles}}{\mathrm{m}^3} = 7.0 \frac{\mathrm{millimoles}}{\mathrm{L}} = 7 \; \mathrm{mM}
\end{equation}

\textcolor{blue}{3.  The authors repeatedly refer to "slithering" mechanisms for rearrangement. I believe this is unfortunate. In the context of the 3SPN model, the "slithering" term was first introduced in Ref. 49 by Sambriski et al, and describes two helical single strands sliding along each other. As noted by Hinckley et al in Ref 18, that mechanism is largely an artifact of the form of the potential in 3SPN.1. The rearrangement mechanism in 3SPN.2 is more of a "defect diffusion", analogous to the "inchworm mechanism" previously identified in the oxDNA model (ref 52 - note, the author list for this reference is wrong). Given this context, it seems unnecessarily confusing to talk about "slithering".}

\textbf{Author Reply:}    (I) We share the reviewer's concerns with using ``slithering'' to describe a process that is more appropriately referred to as ``defect diffusion'', ``inchworm mechanism'', or ``out-of-register diffusion''.

In Section 2.1.2 on \rood{p.~8}, we have modified the text to replace ``slithering'' with ``out-of-register diffusion'':

@@

``An energy disconnectivity graph-based approach was used to interrogate the differences in hybridization rates and mechanisms between GGGGGG and GCGCGC hexamers to reveal strong deviations from ``all-or-nothing'' behaviors and the importance of zippering and \rood{out-of-register diffusion mechanisms.~\citep{Xiao2019}''}

The only other place we use the term is in Section 3.4 on \rood{p.~27} where we refer to work conducted in Ref.~\citep{Xiao2019}. Here we choose to maintain the term ``slithering'' to maintain consistency with the mechanism nomenclature adopted in this paper, but place it in quotation marks to indicate that we are quoting from that paper and it is not a term that we would necessarily favor in the context of our own work.

@@

\rood{``Out-of-register states for 5$^\prime$-GCGCGC-3$^\prime$ hexamers were identified as deep kinetic traps along the hybridization pathway and ``slithering'' through these states did not provide a significant hybridization pathway compared to an alternative ``zippering'' mechanism. (In contrast, out-of-register slithering and in-register zippering served as two parallel pathways for hybridization of 5$^\prime$-GGGGGG-3$^\prime$.)'' }

(II) We also appreciate the reviewer pointing out the incorrect author list in ref 52, and we have fixed the issue.

@@

\textcolor{blue}{4. P15: I am a bit surprised that a separation of 1.3nm between the central base pairs is enough to guarantee strand separation, given that frayed end can separate by 1.3nm when the core is intact. Can the authors verify that all their duplexes had truly separated at this point?}

\textbf{Author Reply:}   This a good point of clarification. We motivated our choice of this distance as that beyond which the strands are effectively non-interacting, but also verified that our estimates of $k_d^{slow}$ are robust to choices of this cutoff in the range 0.8--2.5 nm. We illustrate this in \textbf{Figure R1} below where we show that our extracted values of $k_d^{slow}$ are robust in changing the cutoff from 1.3 nm to 2.0 nm.

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=\textwidth]{./revision_figures/slow_response_compare_cutoff.png}
        \caption*{\textbf{Figure R1.} Comparing the experimental slow response when increasing the cutoff parameter from 1.3 nm to 2.0 nm}
        \label{fig:compare_cutoffs}
	\end{center}
\end{figure}

We have modified the relevant text in Section 3.1 on \rood{p.~15} to read:

``First, we tracked the slow response corresponding to duplex dissociation in our simulations by compiling the distribution of times at which both of the central base pairs first separate to a distance of 1.3 nm starting from an initial fully hybridized duplex. This cutoff was selected as the distance beyond with the strands are effectively non-interacting and define the dissociated state. \rood{We verified that our results were robust to choices of this cutoff over the range 0.8-2.5 nm.}''

@@

\textcolor{blue}{5. The authors refer to "exponential" increase in Fig. 1. However, it's hard to see that an increase is exponential by eye unless the data is plotted on a log scale.}

\textbf{Author Reply:}   We agree that a log-linear plot would better expose these trends. We have replotted Fig.~1 and Fig.~S\rood{4} on log-linear axes and reproduce these below.

\noter{Insert here revised Fig. 1 and Fig. S4 complete with new captions.}

@@

\textcolor{blue}{6. Do the authors have any data on the scaling of k\_hybridization with T? In experiment, rates appear to increase with temperature (albeit with a much weaker dependence than de-hybridization rates). Simple "nucleation and zipper" models would predict the opposite, as does the oxDNA model (see ref 52).}

### 2021-07-21 cover_letter/response_letter.tex

\textbf{Author Reply:}   This is an interesting question and are, in fact, considering embarking upon some additional work in this vein. We do not, however, have any data on this scaling at the current time. Rigorous engagement of this question would require us to fit MSMs at a variety of temperatures in order to compare the pathways, mechanisms, and rates for hybridization and dehybridization. In the present work, we have only constructed MSMs at the melting temperature of each sequence and the collection of sufficient data to parameterize MSMs at other temperatures would require both (i) a substantial investment of time to collect the $\sim$1 ms of data required to fit robust MSMs and (ii) the use of biasing techniques in order to adequately sample hybridization (dehybridization) events at large temperature excursions above (below) the melting temperature. As such, we do not have the necessary data to engage this question in the present work, but would like to explore it in the future.

Reviewer \#2 raised a related point in their Comment \#2 and we have made the following modification to the terminal paragraph of the Conclusions on \rood{p.~35} observing that this may be an interesting follow-up study:

``Going forward, we will extend this work to discern more general trends in sequence-dependent hybridization/dehybridization for a wider range of oligomer sequences and motivate strategies for experimental comparisons. \rood{We also suggest that the MSM approach followed in this work, possible coupled with biased sampling and reweighting techniques \noter{Add refs to TRAM / DHAM or whatever it is that PyEmma uses by default}, may be well-suited to exposing changes in the hybridization/dehybridization mechanism and kinetics as a function of temperature.}''

%We are interested in temperature and sequence affects on  k\_hybridization and exploring a some future work in this vein. Indeed, it is possible to use similar methods to investigate the hybridization process, however these calculation require much greater computational expense in order to gather adequate statistics. As this section served more as a validation for the model’s ability recapitulate experimental responses, we do not think this work is in scope of this work. Furthermore, the k\_hybridization values gathered by T-jump are significantly less reliable than k\_dissociation as the system cools back down and timescales approach the millisecond regime. \noteg{check with Brennan on this}

%\noteg{can cite some experimental work here, but the trend they describe here is not really correct}

@@

\textcolor{blue}{7. The authors could explain a bit more carefully (perhaps in the SI) how they handle the multiple nearest-neighbour states that correspond to their microstates (eg. F4 clearly involves more than one NN configuration).}

\textbf{Author Reply:}   \noter{I think probably we do not want to make too heavy weather of this comment. I think that all is required here is a clearer statement in regarding how the averaging operation is done in collapsing the NN predictions for microstates into a single number for the macrostate. There is no need, I think, for anything more that this and no need to focus specifically on F4 as this is just raised as an example and the revised text should apply equally to all macrostates.}

We have modified the text in the ``Nearest neighbor model of duplex thermodynamics'' text in the Supporting Information to better describe for the reader how we performed this calculation. Specifically, we have appended the following text at the end of this discussion:

``\rood{In applying the NN models to each macrostate, we made the simplifying assumption that the ensemble of microstates constituting each macrostate could be represented by a single pattern of Watson-Crick base pairing that are schematically illustrated in Fig.~2a.}''

%\noteg{Brennan working on this}  NN calculations only explicitly considered configurations where four base pairs are entirely unbound on either side, and the rest are fully bound. The analysis below adds configurations with partial fraying on one side to better treat the diversity of microstates present in F4 (Fig.~\ref{fig:F4_microstates}).

%\begin{figure}[ht!]
%	\begin{center}
%        \includegraphics[width=\textwidth]{./revision_figures/F4_microstates.png}
%        \caption{Microstates distribution within the F4 macrostate. Fraying in the permutable up/down direction is symmetrized such that all frayed ends are on the left and all bound ends on are the right. Panel B shows the percentage of distances greater than a 1.3 nm cutoff, where 21\% of states have at least one base paired frayed and 11\% of states have at least two base pairs frayed. \noteg{Brennan calculating how the inclusion of these states affects the estimated free energy.}}
%        \label{fig:F4_microstates}
%	\end{center}
%\end{figure}

@@

\textcolor{blue}{8. I don't understand how the population of F4 could be >50\% that of H for the GC-core system, when the free energy of F4 is reported as > 5kJ/mol. My back-of-the-envelope maths predicts a ratio of 8:1 for a 5kJ/mol gap. Incidentally, 8:1 is roughly the ratio I got from running the configurations through NUPACK.}

\textbf{Author Reply:}   Our back-of-the-envelope calculations yield a similar result based on the $\Delta F$ values reported in Fig.~3. In brief, assuming:

\begin{itemize}
\item $G_H$ = $G_D$ = 0 (adopting this as the arbitrary zero of free energy and asserting that the sequence is at its melting temperature so the hybridized H and dissociated D states are equiprobable)
\item $G_F$ = 5 kJ/mol = 2 kT
\item $f_H$ + $f_D$ + $f_F$ = 1 (the fractional populations sum to unity)
\item $f_X$ = $\exp (-\beta G_X$) / $Z$ ($Z$ is the partition function that imposes normalization)
\end{itemize}

we can solve this system of equations to yield $f_H$ + $f_D$ = 0.47 and $f_F$ = 0.06, such that $f_H$ / $f_F$ = 7.8. A more rigorous accounting of the state populations by analyzing the MSM predictions yields a slightly lower prediction of $f_F$ = 0.0028 that we report in Fig.~2b. \noter{Mike -- How can we understand this difference? Why is the MSM value so much smaller than expected?}

We believe that the root of this comment may be our choice of a logarithmic y-axis for the bar plots in Fig.~2b that make the bar associated with the F4 state appear nearly 50\% of the height of the H and D bars for GC-core. We made this plotting choice in order to make visible on these bar plots the small state populations associated with the shifted and frayed states and noted the percentage occupancies on the bars themselves. To help eliminate this confusion, we have modified the portion of the Fig.~2 caption referring to this panel to read:

``(b) \textbf{Thermodynamics.} Histograms reporting the number of the $10^7$ total frames within the 1 ms of simulation trajectories observed to occupy each of the seven macrostates, corresponding to our numerical estimates of the equilibrium occupancy probabilities. Uncertainties are calculated across 100 MSMs using a Bayesian MSM estimation are reported for each bar and are very small compared to the total counts. \rood{Values are reported on a log y-axis to make the small populations of the shifted and frayed states visible.}''

@@

\textcolor{blue}{9. Is the MSM model explicitly constrained to obey detailed balance? The numbers that come out seem consistent with detailed balance - forward and backward trajectories at the same T essentially look the same (they're equally likely to go through intermediates). This should be a feature of the MSM, but it's unclear to me whether it's hard-coded or just arises naturally in the coarse-graining.}

\textbf{Author Reply:}    The reviewer raises a good point regarding the assumptions made during MSM construction. We do indeed use the reversible Bayesian MSM framework described here \url{http://www.emma-project.org/latest/api/generated/pyemma.msm.BayesianMSM.html} which enforces detailed balance. This is an important point which we have now made explicit by modifying the text in Section 2.1.2 on \rood{p.~11} to read:

``The 200 microstates comprising each system were coarsened into our terminal macrostate MSM. \rood{Since we are interested in the construction of models of the equilibrium kinetics, we explicitly enforce detailed balance in the construction of the MSM. \noter{Maybe add a reference here. Possibly https://doi.org/10.33011/livecoms.1.1.5965}}''

and the text in Section 3.4 on \rood{p.~24} to read:

@@

``In addition to thermodynamic stabilities, the macrostate MSM also furnishes quantitative and interpretable predictions of hybridization and dehybridization pathways and mechanisms. \rood{It should be noted that because we are using a reversible MSM framework, detailed balance is enforced by construction.''}

\textcolor{blue}{Relatedly, some of the discussion of the hybridization and de-hybridization pathways in non-repetitive sequences seems to ignore this fact. In these equilibrium simulations, hybridization trajectories played in reverse should be statistically indistinguishable from de-hybridization trajectories. I don't think anything that's said is wrong, per se, but it might be worth considering how the contrast between "nucleation and zippering" and "fraying-peeling" is consistent with the symmetry in the pathways.}

\textbf{Author Reply:}  This is a nice observation, and we would do well to note this symmetry for the reader. We have added a sentence to this effect in Section 3.6 on \rood{p.~30}:

\noter{Do you agree with all of this?}
``In dehybridization, we observe fraying \rood{of one half of the duplex with the strands remaining associated until the loss of the final A:T pairs and ultimately the last G:C pair}. Qualitatively, we observed some short-lived states composed of two to four native WC base pair contacts immediately before full dissociation occurs, but, in contrast to the F4 state we observe in GC-core, these conformations do not constitute a metastable state within our MSM nor do they tend to reform intact duplexes. These dissociation dynamics are consistent with a ``fraying-peeling'' dehybridization mechanism.\citep{Wong2008TheSimulations, Perez2010Real-timeUnfolding, Zgarbova2014BaseRNA} \rood{We observe that the principle of microscopic reversibility for a molecular system at thermodynamic equilibrium imposes the conditions of detailed balance and symmetry of the classical equations of motion under time reversibility \noter{Add ref: 10.1021/bi800118b}. A consequence of these considerations is that if our system is indeed at equilibrium, then the ``fraying-peeling'' dehybridization mechanism can be considered a reversal of the ``nucleation-zippering'' hybridization pathway since the time-reversed simulation trajectories represent an equally valid sampling of the system dynamics. This is indeed borne out by our mechanistic observations for GC-mix wherein the early stages of ``nucleation-zippering'' proceed by the formation of one G:C contact followed by one or more additional A:T pairs and the late stages by the formation of all remaining WC pairs, which we compare with the early stages ``fraying-peeling'' wherein one half of the duplex frays and the late stages wherein dissociation finally completes by the dissolution of the last few A:T contacts and the final G:C pair.}''

@@

\textcolor{blue}{10. I think some of the characterisation of the "all or nothing" picture is a bit misleading. The two state model of duplex formation is really a thermodynamic model, based on the assumption that the ensemble of hybridized and de-hybridized states look very different and are separated by a large barrier. This model is perfectly consistent with the transition between the two ensembles involving a complex network of metastable states. Indeed, as the authors note on line 3 of p4, a number of previous authors have considered what such states might look like, particularly for repetitive sequences. The authors' primary contribution here is not proposing that these states exist for the first time, but trying to tie together evidence from simulation and experiment to describe them.}

%\noteg{The ``all-or-nothing'' comment is an important point. Brennan had cautioned using this term necessarily synonym for ``two-state'' and the reviewer seems to be making the same point. There are 11 instances of ``all-or-nothing'' in the main text, and it is often used in conjunction with ``two-state all-or-nothing.'' I think we may want to consider removing most of them, but we should probably word this carefully.}

\textbf{Author Reply:}   We thank the reviewer for this important observation and illuminating our need to better clarify this potentially ambiguous concept. We have modified the text in Section 3.6 on \rood{p.~29} to read:

\noter{What do you think of this? Does this answer the question?}
``The GC-mix sequence is the only one of the four sequences studied that exhibits simple two-state ``all-or-nothing'' behavior.~\citep{Xiao2019, Araque2016LatticeCooperativity, Sikora2013ModelingIntermediates, Sanstead2016} \rood{This characterization is intended to reflect that the mechanistic observation that} GC-mix MSM comprises just two states, the hybridized H and dehybridized D (Fig.~\ref{fig:allseq_table}c), indicating that association and dissociation of the strands proceeds directly without passing through any metastable intermediate states resolvable under the $\tau$=1.2 ns lag time of our MSM. \rood{(We note that ``all-or-nothing'' can also be used, somewhat ambiguously, to refer to a two-state thermodynamic model where the hybridized and dehybridized are separated by a large free energy barrier but that the transitions between them may proceed through a network of metastable states.)}''

\textcolor{blue}{11. Related to the above, the references on line 3 of p4 should probably also include Ref 52, where a similar analysis of metastable states in a repetitive sequence is performed (for a different coarse-grained model) and [Flamm, W., Fontana,J.M., Hofacker, I.L. and Schuster,P. (2000) RNA folding at elementary step resolution. RNA, 6, 325-338] where a defect diffusion mechanism is proposed.
}

\textbf{Author Reply:}   We have added this reference.

@@

\textcolor{blue}{12. The reasoning behind why the broader relaxation spectrum is indicative of misaligned states was not that clear to me. Why would "elevated" fraying cause a broader response? Wouldn't it also, then, be broader for CG-core, which has elevated fraying? It seems to me that the argument is more along the lines of the heterogeneity of the configurations leading to heterogeneous dynamics, and hence a broader relaxation profile. But I am no expert on this technique.}

\textbf{Author Reply:}   The reviewer makes a good comment on our choice of language used to described the stretched AT-all response. It is indeed the heterogeneity of the configurations in the initial ensemble, and the fraying dynamics exhibited in each of these states, that is the origin of the broadening of the relaxation profile, but our wording here was insufficiently clear. We believe that the reviewer's apt suggestion ``heterogeneity of configurations leads to heterogeneous dynamics'' expresses this idea much more clearly and we have integrated it into our edits of Section 3.7 on \rood{p.~32}:

``Nevertheless, the presence of these out-of-register shifted states in the low-temperature equilibrium ensemble prior to the T-jump step may be observable via their influence on the fast T-jump IR response attributable to fraying. Specifically, we hypothesize that the dangling ends and inert tails present in the out-of-register shifted states should promote an \rood{elevated fraying response over the course of the relaxation that is distinct from that of in-register fraying. This heterogeneity of configurations should lead to heterogeneous dynamics, manifested in the observation of a more stretched relaxation over experimental time scales of 70-100 ns.} Analysis of the MSM equilibrium distributions reveals 10.0\% of the equilibrium ensemble to reside in out-of-register shifted states 3S2, 5S2, 3S4, and 5S4 for AT-all, compared to just 0.23\% for GC-end, and 0\% for GC-core and GC-mix. It is our conjecture that a substantial population of out-of-register shifted states in the pre-T-jump AT-all ensemble should be distinguishable from the GC-end, GC-core, and GC-mix as an elongation of the fast relaxation response associated with terminal base fraying.''

@@

\textcolor{blue}{13. How much of the code is being made available? Ideally, the initialization files for the simulations and the code used to perform the MSM analysis would be publicly available. Additionally, it would be good to publish the raw experimental data and the processing tools used to obtain the final results.}

\textbf{Author Reply:}   As a computationally-focussed paper, we agree that the data and analysis code should be made available to the reader. We have uploaded input files and analysis scripts to Github (\url{https://github.com/mrjoness/DNA_SRV-MSMs}) and \noteg{45 GB} of featurized trajectories to Zenodo (\noteg{Zenodo link}). The links to both of these resources are now noted in the Supporting Information statement on \rood{p.~36}:

\noter{Insert new text added to this section here.}

\noter{We need to add one more sentence on why we will not upload the experimental data. How about this? Check with Brennan and Andrei.} The experimental data used to validate the MD simulations and test the prediction of long-lived metastable shifted states are parts of a larger experimental study of a larger family of DNA oligomers. We are currently preparing additional experimental manuscripts on these systems and would prefer to delay sharing of the experimental data until these publications are complete.

@@

\textcolor{blue}{1. Please include at least the first page for each reference citation for the following incomplete journal reference 48}

Updated.

\textcolor{blue}{2. Please update the publication status of reference 70, if possible.  Providing the DOI is acceptable.}

We are unable to locate any additional citation information beyond the arXiv link.

@@

\textcolor{blue}{3. Please include publisher for the following incomplete book reference: 79.}

Updated.

\textcolor{blue}{4. The TOC graphic should fit in an area no larger than 3.25 in. by 1.75 in. (approx. 8.25 cm by 4.45  cm) and should have adequate resolution and clarity. Confirm that all text is legible at this size.}

Updated and confirmed. \noter{Either version is fine, please confirm it fits in the 3.25 by 1.75 in specification and text is legible.}

\bibliographystyle{achemso}
\bibliography{../references_210514_abv}
\end{document}
